---
title: 'mecfa README'
author: Anthony Raborn
date: '`r Sys.Date()`'
output: github_document
---

`r start.time <- Sys.time()`

# mecfa: Metaheuristic Exploratory/Confirmatory Factor Analysis

An R package which utilizes metaheuristic algorithms to simultaneously find the 
best-fitting model for various factor structures and select which factor 
structure has the highest fit. Currently, simulated annealing is the only 
metaheuristic algorithm in use, but ant colony optimization may be incorporated 
in the future.

# Installation

To install this package, you must have `devtools` installed! It is only found on
GitHub at the moment. 

```{r installing-mecfa, eval=FALSE}
require(devtools)
devtools::install_github("AnthonyRaborn/mecfa") # automatically installs stable branch
```

# Usage

The goal of `mecfa` is to find the best-fitting confirmatory factor analysis
from a range of possible latent factors. 
(If you are unsure of what this means, [here is an article which talks about the basics](https://stats.idre.ucla.edu/spss/seminars/introduction-to-factor-analysis/a-practical-introduction-to-factor-analysis-confirmatory-factor-analysis/)). 
It fits these CFA models using the `lavaan` software.

To start, you need a dataset, some initial models, and the `mecfa` package.

## Data

```{r usage-data}
library(mecfa)

# sample data to use
testData <-
  lavaan::HolzingerSwineford1939
```

The `lavaan` Holzinger-Swineford dataset is a 9-item test of mental ability for 
7th and 8th-grade children.
It also includes other variables which may be of interest for structural equation
modeling, but we will ignore these for now.

The test items are labeled `x1` thru `x9`, so we can use these as our item names
when we create the different initial models.
While the generally-accepted model has 3 factors, we will investigate 2, 3, and 
4 factors for this example.

## Initial Models

```{r usage-initial-models}
factorsNames <-
  c("visual", "textual", "speed", "accuracy")
itemNames <-
  paste0("x", 1:9)
initialModels <-
  createInitialModels(
    maxFactors = factorsNames,
    items = itemNames)
    
cat(
  paste(
    initialModels,
    collapse = "\n\n"
  )
)
```

## Function Specifications

```{r function-specifications}
maxSteps <-
  500
fitStatistic <-
  "rmsea"
maximize <-
  FALSE
lavaan.model.specs <-
  list(
    model.type = "cfa", 
    auto.var = TRUE, 
    estimator = "default",
    ordered = NULL,
    int.ov.free = TRUE, 
    int.lv.free = FALSE, 
    std.lv = TRUE, 
    auto.fix.first = FALSE, 
    auto.fix.single = TRUE, 
    auto.cov.lv.x = TRUE, 
    auto.th = TRUE, 
    auto.delta = TRUE, 
    auto.cov.y = TRUE,
    auto.efa = TRUE,
    missing = "listwise"
    )
maxChanges <-
  2
temperatureFunction <-
  "quadratic"
```


## Execute Function

```{r execute-function}
example <-
  exploratorySA(
    initialModels = initialModels,
    originalData = testData,
    maxSteps = maxSteps,
    fitStatistic = fitStatistic,
    maximize = maximize,
    lavaan.model.specs = lavaan.model.specs,
    temperature = temperatureFunction
  )
```

## Investigate Results

```{r investigate-results}
# show/print method
example

# summary method
summary(example)

# plot method
plot(example)
```

`r round(as.difftime(Sys.time() - start.time, format = "%M:%S"),2)`
